---
title: "Reproducible Research: Peer Assessment 1"
author:CCL
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

```{r}
getwd()
setwd("C:/Users/Administrador/Documents")
activity <- read.csv("activity.csv")
library(ggplot2)
library(lattice)
```

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r}
steps.date <- aggregate(steps ~ date, data=activity, FUN=sum)
ggplot(steps.date, aes(x=steps)) + 
       geom_histogram(colour="black", fill = "red", binwidth = 1000) + 
        labs(title="Histogram of Steps Taken per Day", 
             x = "Number of Steps per Day", y = "Number of times (Count)")         

```

2. Calculate and report the **mean** and **median** total number of steps taken per day

```{r}
mean(steps.date$steps)
median(steps.date$steps)
```

Answer: The mean of total number of steps taken per day is 10766.19 and the median is 10765.


## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5 minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
steps.interval <- aggregate(steps ~ interval, data=activity, FUN=mean)

ggplot(steps.interval, aes(x=interval, y=steps))+
         geom_line(colour="red", size=1.3) + geom_point(colour="red", size=2, shape=21, fill="white")+ labs(title="Average Daily Activity Pattern", x="Interval", y="Number of steps")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
steps.interval$interval[which.max(steps.interval$steps)]
```

Answer: THe 5-minute interval contains the maximum number of steps is the 835TH.


## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)

```{r}
sum(is.na(activity))
```

Answer: The total number of missing values is 2304.

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

I will use the means for the 5-minute intervals.


3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity <- merge(activity, steps.interval, by="interval", suffixes=c("",".y"))
nas <- is.na(activity$steps)
activity$steps[nas] <- activity$steps.y[nas]
activity <- activity[,c(1:3)]
```

4. Make a histogram of the total number of steps taken each day. 

```{r}
steps.date <- aggregate(steps ~ date, data=activity, FUN=sum)
ggplot(steps.date, aes(x=steps)) + 
        geom_histogram(colour="black",fill="red", binwidth=1000) + 
        labs(title="Histogram of Full Steps Taken per Day", 
             x="Number of Steps after populate missing values", 
             y="Count") + 
        theme_bw() 

```

And Calculate and report the **mean** and **median** total number of steps taken per day. 
```{r}
mean(steps.date$steps)
median(steps.date$steps)

```
Mean with missing values:10766.19
Mean without missing values: 10766.19
Median with missing values:10765
Median without missing values: 10766.19

Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

Answer: Yes is different. The impact of the missing data seems rather low and the little increase in the graph peak, not affect significantly our predictions.


## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
library(lubridate)
activity$daytype <- as.factor(ifelse(wday(activity$date, label=TRUE) %in% c("Sat", "Sun"), "weekend", "weekday"))
FT <- split(activity, activity$daytype)

avsteps.weekday <- data.frame(tapply(FT$weekday$steps, FT$weekday$interval, mean))
avsteps.weekend <- data.frame(tapply(FT$weekend$steps, FT$weekend$interval, mean))

names(avsteps.weekday) <- "steps"
names(avsteps.weekend) <- "steps"

avsteps.weekday$interval <- as.integer(row.names(avsteps.weekday))
avsteps.weekend$interval <- as.integer(row.names(avsteps.weekend))

avsteps.weekday$daytype <- factor(rep("weekday", nrow(avsteps.weekday)), levels=c("weekday", "weekend"))
avsteps.weekend$daytype <- factor(rep("weekend", nrow(avsteps.weekend)), levels=c("weekday", "weekend"))


avsteps.bind <- rbind(avsteps.weekday, avsteps.weekend)

```

2. Make a panel plot containing a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}

# create a plot using ggplot
ggplot(avsteps.bind) + aes(x=interval, y=steps) + facet_grid(daytype ~ .) + geom_line() + labs(title="Time Series Plot of Average Number of Steps Per Interval", y="Average number of steps")
```